---
title: "Preferences for income redistribution in unequal contexts: Changes in Latin America between 2008 and 2018"
subtitle: "Descriptive statistics"
author: "Gonzalo Franetovic - Juan Carlos Castillo"
date: "10-09-2024"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

+ Packages

```{r Packages, message=FALSE, warning=FALSE, include=FALSE}

#Install and load packages
pacman::p_load(tidyverse, countrycode, car, zoo, wbstats, Hmisc, stargazer, stats,
       ggthemes, data.table, reshape2, scales, xtable, ggrepel, lme4, here, conflicted,
       texreg, ggpubr)

#Packages conflicts
conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")

```

+ Load data 

```{r Database, echo=FALSE, message=FALSE, warning=FALSE}
#Load database
rm(list = ls())
load(here("input", "proc", "LAPOP_Final.RData"))

```

# Descriptive statistics 

```{r Table 1, echo=FALSE, message=FALSE, warning=FALSE}
#New 
lapop = lapop %>% 
  select(country, countryname, country_wave, country_wave_rec, 
         year, wave, weight1500, wt, redistribution, 
         male_f, age, #married_f, 
         ideology_f, #employment_r, 
         education_r, 
         decile, decile_d10, decile_f, 
         zone, sysconf, 
         gdp, gdp_mean, gdp_dif, gini, gini_mean, gini_dif, welfare,
         corruption, corruption_dif)

#Level 1
lapop_a = lapop %>% 
  select(redistribution, decile, 
         male_f, age, 
         #married_f, 
         ideology_f, sysconf, #employment_r, 
         education_r, zone, decile_d10, decile_f
         ) 

lapop_a = lapop_a %>%
  rename(income = decile) %>% 
  mutate(male = as.numeric(male_f) - 1,
 #        married = as.numeric(married_f) - 1,
         ideology = as.numeric(ideology_f) - 1,
         urban = as.numeric(zone) - 1,
#         employment = as.numeric(employment_r) - 1,
         education = as.numeric(education_r) - 1)


#stargazer(as.data.frame(lapop_a), title="Descriptive Statistics", omit.summary.stat = c("p25", "p75"), type="text")

prop.table(table(lapop_a$male_f))
#prop.table(table(lapop_a$married_f))
#prop.table(table(lapop_a$employment_r))
prop.table(table(lapop_a$ideology_f))
prop.table(table(lapop_a$education_r))
prop.table(table(lapop_a$zone))

#Level 2
country_vars_waves_a= country_vars_waves %>% 
  select(-c(year, welfare))

#stargazer(as.data.frame(country_vars_waves_a), title="Descriptive Statistics", omit.summary.stat = c("p25", "p75"), type="text")

#Level 3
country_vars = country_vars_waves %>% 
  distinct(country, .keep_all=TRUE) %>% 
  select(country, welfare)

prop.table(table(country_vars$welfare))

```

# Support for redistribution by countries. Percentage by category.

```{r Figure 2, echo=FALSE, message=FALSE, warning=FALSE}

#Figure 2: Support for redistribution by countries. Percentage by category.
redist = xtabs(~countryname+redistribution, data = lapop)
redist = prop.table(redist, 1)*100
redist = as.data.frame(redist) %>% 
  arrange(desc(redistribution), Freq) %>% 
  mutate(countryname1 = case_when(countryname=="Bolivia" ~ 1,
                                  countryname=="Peru" ~ 2,
                                  countryname=="Guatemala" ~ 3,
                                  countryname=="Ecuador" ~ 4,
                                  countryname=="Honduras" ~ 5,
                                  countryname=="El Salvador" ~ 6,
                                  countryname=="Mexico" ~ 7,
                                  countryname=="Colombia" ~ 8,
                                  countryname=="Panama" ~  9,
                                  countryname=="Brazil" ~ 10,
                                  countryname=="Paraguay" ~ 11,
                                  countryname=="Uruguay" ~ 12,
                                  countryname=="Chile" ~ 13,
                                  countryname=="Argentina" ~ 14,
                                  countryname=="Costa Rica" ~ 15,
                                  countryname=="Nicaragua" ~ 16,
                                  countryname=="Dominican Republic" ~ 17)) %>% 
  mutate(countryname1 = factor(countryname1,
         levels = c(1:17),
         labels = c("Bolivia", "Peru", "Guatemala", "Ecuador", "Honduras", "El Salvador", "Mexico",
                    "Colombia", "Panama", "Brazil", "Paraguay", "Uruguay", "Chile", "Argentina", 
                    "Costa Rica", "Nicaragua", "Dominican Republic")),
         order = factor(countryname,
         levels = c(1:17),
         labels = c("Bolivia", "Peru", "Guatemala", "Ecuador", "Honduras", "El Salvador", "Mexico",
                    "Colombia", "Panama", "Brazil", "Paraguay", "Uruguay", "Chile", "Argentina", 
                    "Costa Rica", "Nicaragua", "Dominican Republic")))

  figure2 = ggplot(data = redist, 
                   aes(x=reorder(countryname1, order), y = Freq, 
                       fill = factor(redistribution))) + 
    geom_bar(stat = "identity", position = position_stack(reverse = TRUE)) + coord_flip() + labs(x = "", y = "", fill="") +
    guides(fill = guide_legend(reverse = TRUE)) + 
    scale_y_continuous(labels = function(Freq){paste0(Freq, "%")}) +
    theme(panel.background = element_rect(fill = "white"), 
          axis.text=element_text(size=10),
          strip.text=element_text(size=12))  + 
    scale_fill_grey(start=0.8, end=0.2, name = "", 
                  labels = c("1 (Disagree)", "2", "3", "4", "5", "6", "7 (Agree)")) 

figure2
ggsave(here("output", "graphs", "Figure2.jpg"), plot=figure2, height = 4, units="in")

```

# Support for redistribution by country and year. Percentage by category.

```{r Figure 3, echo=FALSE, message=FALSE, warning=FALSE}

#Figure 3: Support for redistribution by country and year
figure3 = ggplot(lapop, aes(x= redistribution,  colour= year)) + 
  geom_line(aes(y = ..prop.., fill = factor(..x..)), stat="count", binwidth = 1, size = 1) +
  xlim(1, 7) +
  facet_wrap(~countryname, ncol=3) +
  labs(x = "Preferencias redistributivas", y = "", fill="Año") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_x_continuous(breaks = c(1:7), labels = c("1 (Desacuerdo)", "2", "3", "4", "5", "6", "7 (De acuerdo)")) +
  theme(
    axis.text=element_text(size=15),
        strip.text=element_text(size=15),
        legend.text=element_text(size=15),
        legend.title=element_text(size=15),
        legend.key.size=unit(0.5,"cm"),
        legend.position = "bottom",
        panel.background = element_rect(fill = "white"),
        axis.text.x = element_text(angle = 90, hjust = 1, size = 10)) +
  theme_hc() + 
  scale_color_grey(start=0.9, end=0.1, name="Año") 

figure3
ggsave(here("Output", "Graphs", "Figure3.jpg"), plot=figure3, height = 10, width = 9, units="in")

```

# Average support for redistribution of income deciles, by country. {no deciles}

```{r Table 2, echo=FALSE, message=FALSE, warning=FALSE}

#Table 2: Average support for redistribution of income deciles, by country
redis_decile = with(lapop, tapply(redistribution, list(countryname, decile_f), mean))
redis_decile = as.data.frame(redis_decile)
redis_decile = as.data.frame(setDT(redis_decile, keep.rownames = TRUE)[])

redis_decile_country = with(lapop, tapply(redistribution, list(countryname), mean))
redis_decile_country = as.data.frame(redis_decile_country)
redis_decile_country = as.data.frame(setDT(redis_decile_country, keep.rownames = TRUE)[])
colnames(redis_decile_country) = c("rn", "Total")

redis_decile = left_join(redis_decile, redis_decile_country) %>% 
  rename(country = rn)

redis_decile

print(xtable(redis_decile), include.rownames=FALSE)

#Statistical differences
countries = c("ARG",
              "BOL",
              "BRA",
              "CHL",
              "COL",
              "CRI",
              "DOM",
              "ECU",
              "GTM",
              "HND",
              "MEX",
              "NIC",
              "PAN",
              "PER",
              "PRY",
              "SLV",
              "URY")

for (i in 1:length(countries)) {
  
  lapop_i = lapop %>% 
      filter(country == countries[i])
  
  countries[i]
  name=paste0("m_income_", countries[i])
  pais= paste0(countries[i])
  
  assign(name, lme4::lmer(redistribution ~ 1 + decile_f + (1 | country_wave), data=lapop_i, weights=wt))

}

screenreg(list(m_income_ARG,
             m_income_BOL,
             m_income_BRA,
             m_income_CHL,
             m_income_COL,
             m_income_CRI,
             m_income_DOM,
             m_income_ECU,
             m_income_GTM),  
          stars = c(0.05), 
          digits = 1,
          custom.model.name = c("ARG",
              "BOL",
              "BRA",
              "CHL",
              "COL",
              "CRI",
              "DOM",
              "ECU",
              "GTM"))

screenreg(list(m_income_HND,
             m_income_MEX,
             m_income_NIC,
             m_income_PAN,
             m_income_PER,
             m_income_PRY,
             m_income_SLV,
             m_income_URY),  
          stars = c(0.05), 
          digits = 1,
          custom.model.name = c("HND",
              "MEX",
              "NIC",
              "PAN",
              "PER",
              "PRY",
              "SLV",
              "URY"))

m_income_ALL = lme4::lmer(redistribution ~ 1 + decile_f + (1 | country_wave), data=lapop, weights=wt)

screenreg(m_income_ALL,  
          stars = c(0.05), 
          digits = 1,
          custom.model.name = c("ALL"))
           
```

# Average support for redistribution and GINI, by country and year.

```{r Figure 4, echo=FALSE, message=FALSE, warning=FALSE}

#Within countries
within = lapop %>%
  mutate(wave = as.numeric(paste0(wave))) %>% 
  group_by(country, wave) %>% 
  summarise_all(mean, na.rm=TRUE)

#Figure 4: Average support for redistribution and inequality, by country and year

figure4 <- ggplot(within, 
            aes(x = gini, y = redistribution, label = country)) +
  geom_point(size = 1.5, alpha = 0.8) + 
  stat_smooth(size = 1, method = "lm", se = FALSE, fullrange = TRUE, colour = "black") +   
  geom_text_repel(aes(label = country), size = 2.5, show.legend = FALSE) +
  labs(x = "Desigualdad (Gini)", y =  "Preferencias redistributivas",
       caption = "Fuente: LAPOP 2008-2023; estadísticas descriptivas ponderadas; 
las cifras informan de las medias de los países en preferencias redistributivas según índice de GINI.") + 
  scale_color_grey(start = 0.8, end = 0.2, name = "Año") +  
  scale_y_continuous("Media de preferencias redistributivas", limits = c(1, 7), 
                     breaks = c(1, 2, 3, 4, 5, 6, 7)) +
  scale_x_continuous("GINI", limits = c(39, 60), 
                     breaks = c(40, 45, 50, 55, 60)) +
  theme(panel.grid.major = element_line(colour = "grey"),
        legend.position = "bottom",
        axis.text = element_text(size = 10),
        strip.text = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.title = element_text(size = 10),
        legend.key.size = unit(1, "cm"),
        panel.background = element_rect(fill = "white")) +  
  facet_wrap(~ wave) +
  stat_cor(aes(label = ..r.label..), method = "pearson", 
           label.x.npc = "right", label.y.npc = "bottom", size = 3, 
           hjust = 1, vjust = -0.7)

figure4
ggsave(here("output", "graphs", "Figure4.jpg"), plot=figure4, height = 5, units="in")

```

# Average support for redistribution and GDP per capita, by country and year.

```{r Figure 5, echo=FALSE, message=FALSE, warning=FALSE}

#Figure 5: Average support for redistribution and economic development, by country and year (facet)
figure5= ggplot(within, 
            aes(x= gdp, y= redistribution, label=country)) +
  geom_point(size = 1.5, alpha = 0.8) + 
  stat_smooth(size = 1, method = "lm", se = FALSE, fullrange = TRUE, colour = "black") +   
  geom_text_repel(aes(label=country), size=2.5, show.legend = FALSE) +
  labs(x = "Inequality", y =  "Support for redistribution",
       caption = "Fuente: LAPOP 2008-2023; estadísticas descriptivas ponderadas; 
las cifras informan de las medias de los países en preferencias redistributivas según PIB per capita.") + 
  scale_color_grey(start=0.8, end=0.2, name="Año") +  
  scale_y_continuous( "Media de preferencias redistributivas", limits = c(1,7), 
                     breaks = c(1,2,3,4,5,6,7)) +
  scale_x_continuous("PIB per capita", limits = c(0,16.1), 
                     breaks = c(0,4,8,12,16)) +
  theme(panel.grid.major = element_line(colour = "grey"),
        legend.position="bottom",
        axis.text=element_text(size=10),
        strip.text=element_text(size=10),
        legend.text=element_text(size=10),
        legend.title=element_text(size=10),
        legend.key.size=unit(1,"cm"),
        panel.background = element_rect(fill = "white")) +  
  facet_wrap(~ wave)  +
  stat_cor(aes(label = ..r.label..), method = "pearson", 
           label.x.npc = "right", label.y.npc = "bottom", size = 3, 
           hjust = 1, vjust = -0.7)

figure5
ggsave(here("output", "graphs", "Figure5.jpg"), plot=figure5, height = 5, units="in")

#Detach package conflicted
#detach("package:conflicted", unload = TRUE)

```

# Average support for redistribution and corruption, by country and year.

```{r}
summary(within$corruption)
within$corruption_invert <- -within$corruption  # Invertir los valores de corrupción

figure6 <- ggplot(within, 
       aes(x = corruption_invert, y = redistribution, label = country)) +
  geom_point(size = 1.5, alpha = 0.8) + 
  stat_smooth(size = 1, method = "lm", se = FALSE, fullrange = TRUE, colour = "black") +   
  geom_text_repel(aes(label = country), size = 2.5, show.legend = FALSE, max.overlaps = 20) +
  labs(x = "Corruption", y =  "Support for redistribution",
       caption = "Fuente: LAPOP 2008-2023; estadísticas descriptivas ponderadas; 
las cifras informan de las medias de los países en preferencias redistributivas según la percepción de corrupción de cada país.") +  
  scale_color_grey(start = 0.8, end = 0.2, name = "Wave") +  
  scale_y_continuous("Media de preferencias redistributivas", limits = c(1,7), 
                     breaks = c(1,2,3,4,5,6,7)) +
  scale_x_continuous("Corrupción", 
                     limits = c(-2.5, 2.5), 
                     breaks = c(-2.5, 0, 2.5), 
                     labels = c("Baja", "Media", "Alta")) +
  theme(panel.grid.major = element_line(colour = "grey"),
        legend.position = "bottom",
        axis.text = element_text(size = 10),
        strip.text = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.title = element_text(size = 10),
        legend.key.size = unit(1, "cm"),
        panel.background = element_rect(fill = "white")) +  
  facet_wrap(~ wave) +
  stat_cor(aes(label = ..r.label..), method = "pearson", 
           label.x.npc = "right", label.y.npc = "bottom", size = 3, 
           hjust = 1, vjust = -0.7)

figure6
ggsave(here("output", "graphs", "Figure6.jpg"), plot=figure6, height = 5, units="in")
```


```{r}
#Figure 7: Average support for redistribution and trust in institutions, by country and year (facet)
figure7= ggplot(within, 
            aes(x= sysconf, y= redistribution, label=country)) +
  geom_point(size = 1.5, alpha = 0.8) + 
  stat_smooth(size = 1, method = "lm", se = FALSE, fullrange = TRUE, colour = "black") +   
  geom_text_repel(aes(label=country), size=2.5, show.legend = FALSE, max.overlaps = 30) +
  labs(x = "Inequality", y =  "Support for redistribution",
       caption = "Fuente: LAPOP 2008-2023; estadísticas descriptivas ponderadas; 
las cifras informan de las medias de los países en preferencias redistributivas según media de confianza en instituciones.") +  
  scale_color_grey(start=0.8, end=0.2, name="Wave") +  
  scale_y_continuous( "Media de preferencias redistributivas", limits = c(1,7), 
                     breaks = c(1,2,3,4,5,6,7)) +
    scale_x_continuous("Confianza en instituciones", limits = c(1,7), 
                     breaks = c(1,2,3,4,5,6,7),
                     labels=c("1 
Baja", "2", "3", "4", "5", "6", "7 
Alta")) +
  theme(panel.grid.major = element_line(colour = "grey"),
        legend.position="bottom",
        axis.text=element_text(size=7.9),
        strip.text=element_text(size=10),
        legend.text=element_text(size=10),
        legend.title=element_text(size=10),
        legend.key.size=unit(1,"cm"),
        panel.background = element_rect(fill = "white")) +  
  facet_wrap(~ wave)  +
  stat_cor(aes(label = ..r.label..), method = "pearson", 
           label.x.npc = "right", label.y.npc = "bottom", size = 3, 
           hjust = 1, vjust = -0.7)

figure7
ggsave(here("output", "graphs", "Figure7.jpg"), plot=figure7, height = 5, units="in")
```

# Average support for redistribution and educational level, by country and year.

```{r}
figure8 <- ggplot(lapop, aes(x = education_r, y = redistribution)) +
  geom_violin(trim = FALSE, fill = "lightblue", color = "black", alpha = 0.7) +  # Gráfico de violín
 # geom_jitter(width = 0.2, size = 1.5, alpha = 0.8) +  # Opcional: superponer puntos para mayor detalle
  labs(x = "Nivel educacional", y = "Media de preferencias redistributivas",
       caption = "Fuente: LAPOP 2008-2023; estadísticas descriptivas ponderadas; 
las cifras informan de las preferencias redistributivas para todos los países según nivel educacional.") + 
  scale_y_continuous("Preferencias redistributivas", limits = c(1,7), 
                     breaks = c(1:7)) +
  scale_x_discrete(labels=c("Primaria", "Secundaria", "Terciaria"))+
  theme(panel.grid.major = element_line(colour = "grey"),
        legend.position = "bottom",
        axis.text = element_text(size = 10),
        strip.text = element_text(size = 10),
        panel.background = element_rect(fill = "white")) +
  facet_wrap(~ wave)  # Facets por año


figure8
ggsave(here("output", "graphs", "Figure8.jpg"), plot=figure8, height = 5, units="in")
```

# Average support for redistribution and ideology, by country and year.

```{r}
figure9 <- ggplot(lapop, aes(x = ideology_f, y = redistribution)) +
  geom_violin(trim = FALSE, fill = "lightblue", color = "black", alpha = 0.7) +
#  geom_text_repel(aes(label = country), size = 2.5, color = "black", show.legend = FALSE) +  # Agrega nombres de país en negro
  labs(x = "Ideología", y = "Media de preferencias redistributivas",
       caption = "Fuente: LAPOP 2008-2023; estadísticas descriptivas ponderadas; 
las cifras informan de las preferencias redistributivas para todos los países según posición política.") + 
  scale_y_continuous("Preferencias redistributivas", limits = c(1,7), 
                     breaks = c(1:7)) +
  scale_x_discrete(labels=c("Derecha", "Centro", "Izquierda", "No
declarado"))+
  theme(panel.grid.major = element_line(colour = "grey"),
        legend.position = "bottom",
        axis.text = element_text(size = 10),
        strip.text = element_text(size = 10),
        panel.background = element_rect(fill = "white")) +
  facet_wrap(~ wave)  # Facetas por año

figure9

ggsave(here("output", "graphs", "Figure9.jpg"), plot=figure9, height = 5, units="in")
```

# Average support for redistribution and age, by country and year.

```{r}
#Figure 10: Average support for redistribution and trust in institutions, by country and year (facet)
figure10= ggplot(within, 
            aes(x= age, y= redistribution)) +
  geom_point(size = 1.5, alpha = 0.8) + 
  stat_smooth(size = 1, method = "lm", se = FALSE, fullrange = TRUE, colour = "black") +   
  geom_text_repel(aes(label=country), size=2.5, show.legend = FALSE) +
  labs(x = "Media de edad", y =  "Support for redistribution",
       caption = "Fuente: LAPOP 2008-2023; estadísticas descriptivas ponderadas; 
las cifras informan de las medias de preferencias redistributivas por país según la media de edad.") + 
  scale_color_grey(start=0.8, end=0.2, name="Wave") +  
  scale_y_continuous( "Media de preferencias redistributivas", limits = c(1,7), 
                     breaks = c(1:7)) +
#  scale_x_continuous("Income (decile", limits = c(1,7), 
#                     breaks = c(2,4,6,8,10)) +
  theme(panel.grid.major = element_line(colour = "grey"),
        legend.position="bottom",
        axis.text=element_text(size=10),
        strip.text=element_text(size=10),
        legend.text=element_text(size=10),
        legend.title=element_text(size=10),
        legend.key.size=unit(1,"cm"),
        panel.background = element_rect(fill = "white")) +  
  facet_wrap(~ wave)   +
  stat_cor(aes(label = ..r.label..), method = "pearson", 
           label.x.npc = "right", label.y.npc = "bottom", size = 3, 
           hjust = 1, vjust = -0.7)

figure10
ggsave(here("output", "graphs", "Figure10.jpg"), plot=figure10, height = 5, units="in")
```








